\Huge Literature Search Made Easy: Mining Online Archives for Research Topic Trends

### 2013-11-20 article_1.tex
Resolved a conflict on milestone report

We used abstracts of the first 6000 submissions to the condensed matter section as our corpus. A term matrix generator (TMG) \url{http://scgroup20.ceid.upatras.gr:8000/tmg/} with stop words filter is used to generate a dictionary containing all terms that appeared in at least 5 documents and a term matrix whose (i, j)-th element n denotes that the j-th word in our dictionary occurred n times in the i-th document. The resulting matrix had 19327 terms and was fed thru an off the shelf C-package "hlda-c" \cite{1} implementing hLDA.\\
A depth 4 tree generated by "hlda-c" looks like:\newline
			\begin{figure}[!ht]
				\centerline{\includegraphics[scale = 0.35]{tree10000.png}}
				\caption{Unsupervised Hierarchical Classification of First 10000 Submissions to Condensed Matter Physics Section}
			\end{figure}
\newline
Despite being completely unsupervised, the result is actually surprisingly good. The three numbers in square brackets before each line are tree level, number of terms and number of documents in this branch. The top node (0th - level ) contains the function words. In the next level, the documents are classified into three branches: first class containing 5539 documents, second class containing 417 documents and the third class containing 44 documents. Eventhough the dataset seem to very skewed, with the largest class containing 100 times more documents than the smallest class, the classification captures sub fields in condensed matter physics very nicely: the first class is still very large, so its keywords are still pretty much function words, but if we go a level down to level 2, we can see the first branch contains mostly theoretical studies on correlated systems such as superconductors, antiferromagnet, quantum Hall and Bose Eistein condensation; the second branch contains experimental papers on material growth and transport measurements; Second class on level 1 is topics on studies in the intersection of condensed matter physics and chemistry, namely material synthesis; third branch on level 1 contains topics on computational physics -- using computational methods to calculate physical and chemical properties of materials. The hierarchical clustering not only gives us a very good idea what the sub fields are, but also ranked them according to their popularity, which agrees well with our knowledge of the field.

\section*{Milestone: Progress So Far}
So far we have done the following tasks.

\section*{Milestone: Next Step}

@@

>>>>

We used abstracts of the first 6000 submissions to the condensed matter section as our corpus. A term matrix generator (TMG) \cite{5} with stop words filter is used to generate a dictionary containing all terms that appeared in at least 5 documents and a term matrix whose (i, j)-th element n denotes that the j-th word in our dictionary occurred n times in the i-th document. The resulting matrix had 19327 terms and was fed thru an off the shelf C-package "hlda-c" \cite{1} implementing hLDA.\\
A depth 4 tree generated by "hlda-c" looks like:\newline
			\begin{figure}[!ht]
				\centerline{\includegraphics[scale = 0.35]{tree10000.png}}
				\caption{Unsupervised Hierarchical Classification of First 10000 Submissions to Condensed Matter Physics Section}
			\end{figure}
\newline
Despite being completely unsupervised, the result is actually surprisingly good. The three numbers in square brackets before each line are tree level, number of terms and number of documents in this branch. The top node (0th - level ) contains the function words. In the next level, the documents are classified into three branches: first class containing 5539 documents, second class containing 417 documents and the third class containing 44 documents. Eventhough the dataset seem to very skewed, with the largest class containing 100 times more documents than the smallest class, the classification captures sub fields in condensed matter physics very nicely: the first class is still very large, so its keywords are still pretty much function words, but if we go a level down to level 2, we can see the first branch contains mostly theoretical studies on correlated systems such as superconductors, antiferromagnet, quantum Hall and Bose Eistein condensation; the second branch contains experimental papers on material growth and transport measurements; Second class on level 1 is topics on studies in the intersection of condensed matter physics and chemistry, namely material synthesis; third branch on level 1 contains topics on computational physics -- using computational methods to calculate physical and chemical properties of materials. The hierarchical clustering not only gives us a very good idea what the sub fields are, but also ranked them according to their popularity, which agrees well with our knowledge of the field.

\section*{Milestone: Progress So Far}
So far we have done the following tasks.
\begin{enumerate}
\item Crawled paper titles and abstracts from arXiv
\item Indexed data into term-frequency matrix
\item Transformed data to the LDA input format
\item Clustered with LDA and hLDA
\item Obtained and analyzed initial results
\end{enumerate}

\section*{Milestone: Next Step}
For the next step, we have following plans.
\begin{enumerate}
\item Cluster data with DTM and analyze the evolution of research topics
\item Train and compare the results of sLDA
\item Do experiments with data from different sub domains (e.g. high energy physics, astro physics) and macro domains (e.g. computer science, mathematics)
\end{enumerate}

@@

 \bibitem[5] {5}
  D. Zeimpekis and E. Gallopoulos
  \emph{TMG: A MATLAB-based term-document Matrix Constructor for Text Collections}.
  Technical Report, Computer Engineering and Information Department, University of Patras
  Dec 2003
  http://scgroup20.ceid.upatras.gr:8000/tmg/
